---
title: "Machine Learning Introduction"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Machine Learning with caret

The caret package (short for classification and regression training) contains functions to streamline the model training process for complex regression and classification problems. The package utilizes a number of R packages but tries not to load them all at package start-up1. The package “suggests” field includes 27 packages. caret loads packages as needed and assumes that they are installed. Install caret using `install.packages("caret", dependencies = c("Depends", "Suggests"))` to ensure that all the needed packages are installed.
data
[The caret Package](http://topepo.github.io/caret/index.html) has several functions that attempt to streamline the model building and evaluation process, as well as feature selection and other techniques.

### Exercise 

The R caret package will make your modeling life easier – guaranteed. caret allows you to test out different models with very little change to your code and throws in near-automatic cross validation-bootstrapping and parameter tuning for free.

For example, below we show two nearly identical lines of code. Yet they run entirely different models. In the first, method = "lm" tells caret to run a traditional linear regression model. In the second line method = "rf" tells caret to run a random forest model using the same data. One tiny syntax change and you run an entirely new type of model.

`library(caret)`

*In R, the `annual_pm ~ .` means use annual_pm as the model response*
*and use all other variables as predictors*

`lm1 <- train(annual_pm ~ ., data = air, method = "lm")`
`rf1 <- train(annual_pm ~ ., data = air, method = "rf")`


```{r two-plus-two, exercise=TRUE, exercise.eval=TRUE}
library(caret)
lm1 <- train(annual_pm ~ ., data = air, method = "lm")
rf1 <- train(annual_pm ~ ., data = air, method = "rf")
```

### Exercise with Code

One of the most powerful aspects of the caret package is the consistent modeling syntax.
By simply changing the method argument, you can easily cycle between, for example, 
running a linear model, a gradient boosting machine model and a LASSO model. In total, 
there are 233 different models available in caret. 

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

