---
title: "Exercises - Data Visualization"
output: html_notebook
---
```{r}
library(tidyverse)
```


# Exercises for the Data Visualization Presentation

1. Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```

2. How many rows are in mpg? How many columns?

```{r}
mpg
```

3. What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

4. Make a scatterplot of hwy vs cyl.

```{r}
ggplot(mpg) +
     geom_point(aes(x = hwy, y = cyl))
```

5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(mpg) +
     geom_point(aes(x = class, y = drv))
```

6. What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

Since you are setting the color manually, it needs to be set outside of the 
aes() mapping.

7. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the
documentation for the dataset). How can you see this information when you run mpg?

```{r}
?mpg
str(mpg)
```

8. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for
categorical vs. continuous variables?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl, size = cty, shape = year ))
```

9. What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = trans, size = trans, shape = trans ))
```

10. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y = hwy, stroke = cyl))
```

For shapes that have a border (like 21), you can colour the inside and
outside separately. Use the stroke aesthetic to modify the width of the
border

11. What happens if you map an aesthetic to something other than a variable name, like aes(colour =
displ < 5)?

```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

The color is determined by True or False.

12.  What happens if you facet on a continuous variable?

13.  What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean?
    How do they relate to this plot?
    
    ```{r, eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = drv, y = cyl))
    ```

14.  What plots does the following code make? What does `.` do?

    ```{r eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(drv ~ .)
    
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(. ~ cyl)
    ```

15.  Take the first faceted plot in this section:

    ```{r, eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_wrap(~ class, nrow = 2)
    ```
    
    What are the advantages to using faceting instead of the colour aesthetic?
    What are the disadvantages? How might the balance change if you had a 
    larger dataset?
    
16.  Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other
    options control the layout of the individual panels? Why doesn't
    `facet_grid()` have `nrow` and `ncol` argument?

17.  When using `facet_grid()` you should usually put the variable with more
    unique levels in the columns. Why?
    
18.  What geom would you use to draw a line chart? A boxplot? 
    A histogram? An area chart?

19.  Run this code in your head and predict what the output will look like.
    Then, run the code in R and check your predictions.
    
    ```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ```

20.  What does `show.legend = FALSE` do?  What happens if you remove it?  
    Why do you think I used it earlier in the chapter?

21.  What does the `se` argument to `geom_smooth()` do?


22.  Will these two graphs look different? Why/why not?

    ```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth()
    
    ggplot() + 
      geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
    ```

23.  Recreate the R code necessary to generate the following graphs.
    
    ```{r echo = FALSE, fig.width = 3, out.width = "50%", fig.align = "default", message = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(aes(group = drv), se = FALSE) +
      geom_point()
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) +
      geom_smooth(aes(linetype = drv), se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(size = 4, colour = "white") + 
      geom_point(aes(colour = drv))
    ```