---
title: "Getting Started with R"
author: "David Meza"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## Download and Install R

![Tareef Kawaf from RStudio presenting at useR! 2017 Brussels](https://www.youtube.com/watch?v=w0SL8EbLW88&list=PL9HYL-VRX0oSFL5cDLSwDa8jKAa9RG8UR)

R is a free software environment for statistical computing and graphics. It compiles
and runs on a wide variety of UNIX platforms, Windows and MacOS. 

[R Project for Statistical Computing](https://www.r-project.org/)

*Windows*

* Open http://www.r-project.org/ in your browser.
* Click on “CRAN”. You’ll see a list of mirror sites, organized by country.
* Select a site near you.
* Click on “Windows” under “Download and Install R”.
* Click on “base”.
* Click on the link to download the latest version of R (an .exe file).
* When the download completes, double-click the .exe file and answer the usual
questions.

*OS X*

* Open http://www.r-project.org/ in your browser.
* Click on “CRAN”. You’ll see a list of mirror sites, organized by country.
* Select a site near you.
* Click on “MacOS X”.
* Click on the .pkg file for the latest version of R, under “Files:”, to download it.
* When the download completes, double-click the .pkg file and answer the usual
questions.

[RStudio](https://www.rstudio.com/) makes R easier to use. It includes a code editor,
debugging & visualization tools. RStudio is an active member of the R community. 
We believe free and open source data analysis software is a foundation for innovative
and important work in science, education, and industry. The many customers who
value our professional software capabilities help us contribute to this community.

## How to Get Help on a Function

Use help to display the documentation for the function:

`help("hist")`

`?rnorm`

`data()`

`vignette()`

```{r help-functions, exercise = TRUE}

```

Use args for a quick reminder of the function arguments:

*Type this in the console below: `args(tidyr::gather)`*

```{r args, exercise = TRUE}


```

Use **example** to see examples on using a function:

`library(dplyr)`

`example("filter")`

```{r example, exercise = TRUE}

```

Type the function name with out parenthesis to view the function:

Try `heatmap`

```{r heatmap, exercise = TRUE}

```

## Searching the Web for Help

Inside R, use the RSiteSearch function to search by keyword or phrase:
RSiteSearch("key phrase")

In your browser, try using these sites for searching:

[Rseek](http://rseek.org) is a Google custom search that is focused on R-specific websites.

[Stack Overflow](http://stackoverflow.com/) is a searchable Q&A site oriented 
toward programming issues such as data structures, coding, and graphics.

The [Statistical Analysis](http://stats.stackexchange.com/) area on Stack Exchange
is also a searchable Q&A site, but it is oriented more toward statistics than programming.

[RStudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/) make it easy
to learn about and use some of our favorite packages. From time to time, new cheat
sheets will be added to the gallery.

## Helpful Tools

[Cheatsheets](https://www.rstudio.com/resources/cheatsheets/)

Code Snippets are text macros that are used for quickly inserting common snippets
of code.

[Using Code snippets](https://support.rstudio.com/hc/en-us/articles/204463668-Code-Snippets?version=1.0.153&mode=desktop)

[Custom snippets in RStudio](https://maraaverick.rbind.io/2017/09/custom-snippets-in-rstudio-faster-tweet-chunks-for-all/?utm_content=bufferb9ad2&utm_medium=social&utm_source=twitter.com&utm_campaign=buffer) 

RStudio addins were released in early 2016 to provide anyone with the ability to
add "extensions" to RStudio.

[RStudio Addins](https://www.rstudio.com/resources/webinars/understanding-add-ins/)

[Discover and install useful RStudio addins](https://cran.r-project.org/web/packages/addinslist/README.html)

## Reading Files

*Tabular datafiles are quite common. They are text files with a simple format:*

* Each line contains one record.
* Within each record, fields (items) are separated by a one-character delimiter, such
as a space, tab, colon, or comma.
* Each record contains the same number of fields.

Creating a dataframe to store data from a csv file:

```{r readcsv, echo = TRUE, eval = FALSE}
batting1 <- read.csv(file = "Batting.csv")
head(batting1)
```

Creating a dataframe to store data from an xlsx file:



```{r readxl, echo = TRUE, eval = FALSE}
library(readxl)
batting2 <- read_excel(path = "Batting.xlsx")
head(batting2)
```



## Creating A Vector

Use the c(...) operator to construct a vector from given values.
`c(1,1,2,3,5,8,13,21)`

`c(1*pi, 2*pi, 3*pi, 4*pi)`

`c("Everyone", "loves", "stats.")`

`c(TRUE,TRUE,FALSE,TRUE)`

```{r vectors, exercise = TRUE}

```

If the arguments to c(...) are themselves vectors, it flattens them and combines them
into a single vector:

`v1 <- c(1,2,3)`

`v2 <- c(4,5,6)`

`c(v1,v2)`

```{r combine_vector, exercise = TRUE}


```

Vectors cannot contain a mix of data types, such as numbers and strings. If you create
a vector from mixed elements, R will try to accommodate you by converting one of
them:

`v1 <- c(1,2,3)`

`v3 <- c("A","B","C")`

`c(v1,v3)`

```{r mixed_vector,  exercise = TRUE}

```

## Computing Basic Statistics

`x <- c(0,1,1,2,3,5,8,13,21,34)`

`x`

`mean(x)`

`median(x)`

`sd(x)`

`var(x)`

`summary(x)`

```{r basic_stats, exercise = TRUE}

```
 
The cor and cov functions can calculate the correlation and covariance, respectively,
between two vectors:

`x <- c(0,1,1,2,3,5,8,13,21,34)`

`y <- log(x+1)`

`cor(x,y)`

`cov(x,y)`

```{r cor_cov, exercise = TRUE}


```

All these functions are picky about values that are not available (NA). Even one NA
value in the vector argument causes any of these functions to return NA, or even halt
altogether with a cryptic error:

`x <- c(0,1,1,2,3,NA)`

`mean(x)`

`sd(x)`

```{r picky_NA, exercise = TRUE}

```

It’s annoying when R is that cautious, but it is the right thing to do. You must think
carefully about your situation. Does an NA in your data invalidate the statistic? If yes,
then R is doing the right thing. If not, you can override this behavior by setting
na.rm=TRUE, which tells R to ignore the NA values:

`x <- c(0,1,1,2,3,NA)`

`mean(x, na.rm=TRUE)`

`sd(x, na.rm=TRUE)`

```{r, NAremove, exercise = TRUE}


```

Examples of runing basic stats on entire dataframe:

`mtcars`

`var(mtcars)`

`cor(mtcars)`

`summary(mtcars)`

```{r df_stats, exercise = TRUE}


```


## Selecting Dataframe Columns by Position

`dfrm[[n]]`

Returns one column—specifically, the nth column of dfrm.
     
To select one or more columns and package them in a data frame, use the following
sublist expressions:

`dfrm[n]`

Returns a data frame consisting solely of the nth column of dfrm.

`dfrm[c(n1, n2, ..., nk)]`

Returns a data frame built from the columns in positions n1, n2, ..., nk of dfrm.

You can use matrix-style subscripting to select one or more columns:

`dfrm[, n]`

Returns the nth column (assuming that n contains exactly one value).

`dfrm[, c(n1, n2, ..., nk)]`

Returns a data frame built from the columns in positions n1, n2, ..., nk.

`mtcars[c(4, 5, 6)]`

`mtcars[c(4:6)]`

`mtcars[-4]`

`mtcars[c(-4:-6)]`

`mtcars[, 4]`

`mtcars[1:5, 4]`

`mtcars[1:5, 4:6]`

`mtcars[1:5, c(1, 4:6, 10)]`

```{r col_pos, exercise = TRUE}

```

To select a single column, use one of these list expressions:

`dfrm[["name"]]`

Returns one column, the column called name.

`dfrm$name`

Returns the same, but presented in different syntax.

To select one or more columns and package them in a data frame, use these list
expressions:

`dfrm["name"]`

Selects one column and packages it inside a data frame object.

`dfrm[c("name1", "name2", ..., "namek")]`

Selects several columns and packages them in a data frame.

You can use matrix-style subscripting to select one or more columns:

`dfrm[, "name"]`

Returns the named column.

`dfrm[, c("name1", "name2", ..., "namek")]`

Selects several columns and packages them in a data frame.

## Selecting Dataframe Columns by Name

`mtcars[["hp"]]`

`mtcars$hp`

`mtcars["hp"] #dataframe`

`mtcars[c("hp", "wt", "vs", "disp")] #eventhough the are out of order`

`mtcars[, "hp"] #matrix`

`mtcars[, c("hp", "wt", "vs", "disp")]`

`mtcars[, c("hp", "wt", "vs", "disp")] %>% var()`

`mtcars[, c("hp", "wt", "vs", "disp")] %>% cor()`

```{r select_col, exercise = TRUE}


```





