---
title: "Power and Sample Size"
author: "David Meza"
date: "8/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sjstats)
```

# sjstats Package

The sjsats package (https://cran.r-project.org/web/packages/sjstats/sjstats.pdf)

The information below is directly from page 51 of the pdf, detailing the info for the
`smpsize_lmm` function which provides the sample size for linear mixed models.

## Description

Compute an approximated sample size for linear mixed models (two-level-designs), based on powercalculation
for standard design and adjusted for design effect for 2-level-designs.

## Usage

`smpsize_lmm(eff.size, df.n = NULL, power = 0.8, sig.level = 0.05, k, n, icc = 0.05)`

## Arguments

eff.size - Effect size.

df.n -  Optional argument for the degrees of freedom for numerator. See ’Details’.

power - Power of test (1 minus Type II error probability).

sig.level - Significance level (Type I error probability).

k - Number of cluster groups (level-2-unit) in multilevel-design.

n - Optional, number of observations per cluster groups (level-2-unit) in multilevel design.

icc - Expected intraclass correlation coefficient for multilevel-model.

## Details

The sample size calculation is based on a power-calculation for the standard design. If df.n
is not specified, a power-calculation for an unpaired two-sample t-test will be computed (using
pwr.t.test of the pwr-package). If df.n is given, a power-calculation for general linear models
will be computed (using pwr.f2.test of the pwr-package). The sample size of the standard design
is then adjusted for the design effect of two-level-designs **(see deff)**. Thus, the sample size calculation
is appropriate in particular for two-level-designs (see Snijders 2005). Models that additionally
include repeated measures (three-level-designs) may work as well, however, the computed sample
size may be less accurate.

## Value

A list with two values: The number of subjects per cluster, and the total sample size for the linear
mixed model.

## References

*Cohen J. 1988. Statistical power analysis for the behavioral sciences (2nd ed.). Hillsdale,NJ:
Lawrence Erlbaum.*

*Hsieh FY, Lavori PW, Cohen HJ, Feussner JR. 2003. An Overview of Variance Inflation Factors
for Sample-Size Calculation. Evaluation & the Health Professions 26: 239–257. doi: 10.1177/
0163278703255230*

*Snijders TAB. 2005. Power and Sample Size in Multilevel Linear Models. In: Everitt BS, Howell
DC (Hrsg.). Encyclopedia of Statistics in Behavioral Science. Chichester, UK: John Wiley & Sons,
Ltd. doi: 10.1002/0470013192.bsa492*

## Examples

Sample size for multilevel model with 30 cluster groups and a small to
medium effect size (Cohen's d) of 0.3. 27 subjects per cluster and
hence a total sample size of about 802 observations is needed.

```{r}
smpsize_lmm(eff.size = .3, k = 30)
```


Sample size for multilevel model with 20 cluster groups and a medium
to large effect size for linear models of 0.2. Five subjects per cluster and
hence a total sample size of about 107 observations is needed.

```{r}
smpsize_lmm(eff.size = .2, df.n = 5, k = 20, power = .9)
```


# Sample size for study

Based on the examples above and the data you provided, here is the code to determine
sample size:

```{r}
#Sample size for the study
smpsize_lmm(eff.size = .2, df.n = 25, power = 0.8, sig.level = 0.05, k =40)
```

1. df.n is based on the approximate class size
2. power was provided by you
3. I assumed significance level of 0.05
4. K is the total number of groups, 20 treatment and 20 control


eff.size I used from the package example. To determine the study effect size (eff.size)
we can use the `deff()` function. The details are in the pdf.

## Description

Compute the design effect (also called Variance Inflation Factor) for mixed models with two-level
design.

## Usage

deff(n, icc = 0.05)

## Arguments

n - Average number of observations per grouping cluster (i.e. level-2 unit).
icc -  Assumed `intraclass correlation coefficient` for multilevel-model.

Not knowing your thoughts on the icc, I used the default value of 0.05.

```{r}
deff(25)
```

The result shows a effect size of 2.2. If we put this into the `smpsize_lmm()` function, we get:

```{r}
smpsize_lmm(eff.size = 2.2, df.n = 25, power = 0.8, sig.level = 0.05, k =40)
```

I hope this is useful information. You will need to verify I inserted the correct values for the parameters.

Finally here are a couple of other websites I found useful.

### HLM using R

https://www.methodsconsultants.com/tutorial/estimating-hlm-models-using-r-part-1/

### The pwr packge for determining the power and sample size

https://cran.r-project.org/web/packages/pwr/vignettes/pwr-vignette.html

### Overview of the pwr package

http://www.statmethods.net/stats/power.html

