---
title: "Data Transformation"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r setup, include = FALSE}
library(learnr)
library(nycflights13)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
flights <- flights
```

Using the flights dataframe from the nycflights13 package, complete the following exercises.
The flights dataframe has been loaded for you. Below are the variable names.

```{r}
names(flights)
```

###  Filter Exercises

1.  Find all flights that
    1. Had an arrival delay of two or more hours
    
```{r filter1, exercise = TRUE}


```

```{r filter1-solution}
filter(flights, arr_delay >= 120)
#OR
flights %>% 
     filter(arr_delay >= 120)
```
      
    2. Flew to Houston (`IAH` or `HOU`)
    
```{r filter2, exercise = TRUE}


```
 
   
```{r filter2-solution}
filter(flights, dest == "IAH" | dest == "HOU")
```

    3. Were operated by United, American, or Delta
    
```{r filter3, exercise = TRUE}

```

```{r filter3-solution}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

    4. Departed in summer (July, August, and September)
    
```{r filter4, exercise = TRUE}

```

```{r filter4-solution}
filter(flights, month %in% c(7, 8, 9))
```

    5. Arrived more than two hours late, but didn't leave late
    
```{r filter5, exercise = TRUE}

```

```{r filter5-solution}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

    6. Were delayed by at least an hour, but made up over 30 minutes in flight
    
```{r filter6, exercise = TRUE}

```

```{r filter6-hint-1} 
# For the second condition, you will need to test for a difference 
# between depart and arrival time.
```

```{r filter6-hint-2}
filter(flights, dep_delay >= 60, arr_delay - dep_delay <= -30)
```

    7. Departed between midnight and 6am (inclusive)
    
```{r filter7, exercise = TRUE}

```

```{r filter7-solution}
filter(flights, dep_time >= 0 & dep_time <= 0600)
```

2.  Another useful dplyr filtering helper is `between()`. What does it do?
    Can you use it to simplify the code needed to answer the previous 
    challenges?
    
```{r filter8, exercise = TRUE}

```

```{r filter8-solution}
filter(flights, between(dep_time, 0, 0600))
```

3.  How many flights have a missing `dep_time`? What other variables are 
    missing? What might these rows represent?
    **Note: Review the funtions in the dplyr package**

```{r filter9, exercise = TRUE}




```

```{r filter9-solution}
flights %>% count(is.na(dep_time))
```

4.  Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing?
    Why is `FALSE & NA` not missing? Can you figure out the general
    rule?  (`NA * 0` is a tricky counterexample!)
    
### Arrange Exercises

1.  How could you use `arrange()` to sort all missing values to the start?
    (Hint: use `is.na()`).
    
1.  Sort `flights` to find the most delayed flights. Find the flights that
    left earliest.

1.  Sort `flights` to find the fastest flights.

1.  Which flights travelled the longest? Which travelled the shortest?

### Select Exercises

1.  Brainstorm as many ways as possible to select `dep_time`, `dep_delay`,
    `arr_time`, and `arr_delay` from `flights`.
    
1.  What happens if you include the name of a variable multiple times in
    a `select()` call?
  
1.  What does the `one_of()` function do? Why might it be helpful in conjunction
    with this vector?
    
    ```{r}
    vars <- c("year", "month", "day", "dep_delay", "arr_delay")
    ```
    
1.  Does the result of running the following code surprise you?  How do the
    select helpers deal with case by default? How can you change that default?

    ```{r, eval = FALSE}
    select(flights, contains("TIME"))
    ```
    
### Exercises

```{r, eval = FALSE, echo = FALSE}
flights <- flights %>% mutate(
  dep_time = hour * 60 + minute,
  arr_time = (arr_time %/% 100) * 60 + (arr_time %% 100),
  airtime2 = arr_time - dep_time,
  dep_sched = dep_time + dep_delay
)

ggplot(flights, aes(dep_sched)) + geom_histogram(binwidth = 60)
ggplot(flights, aes(dep_sched %% 60)) + geom_histogram(binwidth = 1)
ggplot(flights, aes(air_time - airtime2)) + geom_histogram()
```

1.  Currently `dep_time` and `sched_dep_time` are convenient to look at, but
    hard to compute with because they're not really continuous numbers. 
    Convert them to a more convenient representation of number of minutes
    since midnight.
    
1.  Compare `air_time` with `arr_time - dep_time`. What do you expect to see?
    What do you see? What do you need to do to fix it?
    
1.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you
    expect those three numbers to be related?

1.  Find the 10 most delayed flights using a ranking function. How do you want 
    to handle ties? Carefully read the documentation for `min_rank()`.

1.  What does `1:3 + 1:10` return? Why?

1.  What trigonometric functions does R provide?

